---
title: "EDLD 653 Final"
author: "Chen Su"
format: 
  html:
    embed-resources: true
execute: 
  warning: false
---

### Part 1. API calls to chess.com 

```{r}
#| label: part-1-api-call

library(purrr)
library(httr)
library(jsonlite)

response <- GET('https://api.chess.com/pub/country/US/players')

player_list <- fromJSON(rawToChar(response$content))$players

# extract first 500 usernames
player_usernames <- player_list[1:500]

# create function to pull all the columns for each user
player_data_pull <- function(username) {
  
  info <- GET(paste0('https://api.chess.com/pub/player/', username))
  player_info <- fromJSON(rawToChar(info$content))
  
  stats <- GET(paste0('https://api.chess.com/pub/player/', username, '/stats'))
  player_stats <- fromJSON(rawToChar(stats$content))
  
  # Extract rapid stats if available
  rapid_stats <- player_stats$chess_rapid
  
  list(username = username, 
       name = ifelse(is.null(player_info$name), NA, player_info$name), 
       number_of_followers = player_info$followers, 
       location = player_info$location, 
       last_online = as.POSIXct(
         player_info$last_online, 
         origin = "1970-01-01", 
         tz = "UTC"
         ), 
      rapid_last_rating = ifelse(is.null(rapid_stats$last$rating), NA, rapid_stats$last$rating), 
      rapid_best_rating = ifelse(is.null(rapid_stats$best$rating), NA, rapid_stats$best$rating), 
      rapid_record_win = ifelse(is.null(rapid_stats$record$wins), NA, rapid_stats$record$wins),
      rapid_record_loss = ifelse(is.null(rapid_stats$record$losses), NA, rapid_stats$record$losses), 
      rapid_record_draw = ifelse(is.null(rapid_stats$record$draws), NA, rapid_stats$record$draws)
  )
}

players_df <- map_df(player_usernames, player_data_pull)

str(players_df)

```

*Part 1 Note: Microsoft Copilot was used to assist in troubleshooting the code for rapid stats and date.*



### Part 2. Identify & Remove Duplicate Files
```{r}
#| label: part-2-duplicates

library(dplyr)
library(digest)

files <- list.files("C:/Users/chens/Desktop/Spring 2026/edld653/edld653_final/question_2", 
           full.names = TRUE, recursive = TRUE)

# finding duplicates
file_hashes <- sapply(files, digest, file = TRUE)

hash_table <- data.frame(
  file = files,
  hash = file_hashes
)

originals <- hash_table[!duplicated(hash_table$hash), ]
duplicates <- hash_table[duplicated(hash_table$hash), ]

# create duplicate table
duplicate_table <- inner_join(originals, duplicates, by = "hash") |>
  select(originals = file.x,
         duplicates = file.y)

deleted_duplicates <- vapply(duplicate_table$duplicates, file.remove, logical(1))

# check 
length(list.files("C:/Users/chens/Desktop/Spring 2026/edld653/edld653_final/question_2", 
                  recursive = TRUE))

```

*Part 2 Note: Microsoft Copilot was used to assist in finding duplicates using digest package.*



### Part 3. Multiple Models
```{r}
#| label: models

library(tidyverse)
library(here)
library(lme4)

df <- read.csv(here("benchmarks.csv"))

# recode wave and score, nest by grade (3-5) and subject (reading, math)

grade_subject <- df |>
    mutate(wave = case_when(
                  season == "fall" ~ 0,
                    season == "winter" ~ 1,
                    season == "spring" ~ 2)) |> 
  pivot_longer(cols = c(reading, math),
               names_to = "subject",
               values_to = "score") |>
  group_by(grade, subject) |>
  nest()

# function for true score variance
true_score_variance <- function(data) {
  model <- lme4::lmer(score ~ 1 + wave + (1+wave|sid), REML = FALSE, data = data)
  VarCorr(model)$sid["wave", "wave"]
}

# function for observed score variance
observed_score_variance <- function(data) {
  data %>%
    split(.$sid) %>%
    map(~lm(score ~ 1+ wave, data = .x)) %>%
    map_dbl(~ coef(.x)["wave"]) %>%
    var()
}

slope_table <- grade_subject |>
  mutate(
    true_variance = map_dbl(data, true_score_variance),
    observed_variance  = map_dbl(data, observed_score_variance),
    slope_reliability = true_variance / observed_variance
  ) |>
  select(grade, subject, 
         true_variance, observed_variance, slope_reliability)

print(slope_table)

```

*Part 3 Note: Microsoft Copilot was used to assist in figuring out how to extract the variance of the slope.*



### Part 4. Slope Reliability Function
```{r}
#| label: slope-reliability-function

cal_slope_reliability <- function(model) {
 # extracting model data 
 mod_data <- model.frame(model)
 
 # extracting names of variables used in formulas
 dv_name <- as.character(formula(model))[2]
 cluster_name <- names(lme4::ngrps(model))[1]
 rand_terms <- rownames(lme4::VarCorr(model)[[cluster_name]])
 time_name   <- rand_terms[rand_terms != "(Intercept)"][1]
 
 # extracting true variance
 true_variance <- lme4::VarCorr(model)[[cluster_name]][time_name, time_name]
 
 # calculating observed slope variance
 lm_formula <- as.formula(paste(dv_name, "~ 1 +", time_name))
  
  observed_variance <- mod_data |>
    split(mod_data[[cluster_name]]) |>
    map(~ lm(lm_formula, data = .x)) |>
    map_dbl(~ {
      coef <- coef(.x)
      if (time_name %in% names(coef)) coef[time_name] else NA_real_
      }) |>
    var(na.rm = TRUE)
  
  #calculating slope reliability
  slope_reliability <- true_variance / observed_variance
}


# apply the function

slope_table_function <- grade_subject |>
  mutate(model = map(data, ~lme4::lmer(score ~ 1 + wave + (1+wave|sid), REML = FALSE, data = .x)),
         slope_reliability = map_dbl(model, cal_slope_reliability)) |>
           select(grade, subject, slope_reliability)

print(slope_table_function)

```

*Part 4 Note: Google Gemini was used to assist in troubleshooting how to extract the dv names only, function that skips students without enough data to have a slope, and troubleshooting error from split() for observed variance.*


### Part 5. 95% Confidence Interval (CI) 
```{r}
#| label: ci-function

bootstrap_slope_reliability <- function(data, group_var, time_var, outcome_var, n_boot = 100) { 
  # convert variables to characters 
  group <- deparse1(substitute(group_var)) 
  time <- deparse1(substitute(time_var)) 
  outcome <- deparse1(substitute(outcome_var)) 
  
  # extract student IDs and counts 
  unique_clusters <- unique(data[[group]]) 
  n_clusters <- length(unique_clusters) 
  
  # Create a dynamic formula for lmer 
  formula <- as.formula(paste0( 
    outcome, " ~ 1 + ", time, " + (1 + ", time, " | bootstrap_id)" 
  )) 
  
  boot_estimates <- map_dbl(1:n_boot, function(b) { 
    # Sample student IDs with replacement 
    sampled_id <- base::sample(unique_clusters, size = n_clusters, replace = TRUE) 
    
    # Reconstruct the dataset with dummy IDs 
    boot_data <- map_dfr(seq_along(sampled_id), function(i) { 
      original_id <- sampled_id[i] 
      student_rows <- data[data[[group]] == original_id, ] 
      student_rows[["bootstrap_id"]] <- paste0("boot_", i) 
      return(student_rows) 
    }) 
    
    # Fit model to the bootstrapped data 
    boot_model <- tryCatch({ 
      lmer(formula, data = boot_data, control = lmerControl(calc.derivs = FALSE)) 
    }, error = function(e) return(NULL)) 
    
    if (is.null(boot_model)) return(NA_real_) 
    
    # Calculate reliability on the bootstrap model
    val <- tryCatch({ 
      cal_slope_reliability(boot_model) 
    }, error = function(e) return(NA_real_)) 
    
    return(val) 
  })
  
  # Extract the 95% Confidence Interval 
  ci <- quantile(boot_estimates, probs = c(0.025, 0.975), na.rm = TRUE) 
  
  return(tibble( 
    ci_lower = ci[1], 
    ci_upper = ci[2], 
    boot_mean = mean(boot_estimates, na.rm = TRUE) 
  )) 
}


# applying function
set.seed(123)

slope_reliability_with_ci <- grade_subject |> 
  mutate(
    boot_results = map(data, ~ bootstrap_slope_reliability(
      data = .x,
      group_var = sid,
      time_var = wave,
      outcome_var = score,
      n_boot = 100
    ))
  ) |> 
  unnest(boot_results) |> 
  select(grade, subject, boot_mean, ci_lower, ci_upper)

print(slope_reliability_with_ci)

```

*Part 5 Note: Google Gemini was used to assist in writing the bootstrapping function.*